---
title: "Spotify API"
author: "Jack Swaisgood"
format: html
editor: visual
---

# Introduction

Music is a large part of my daily schedule, and Spotify allows me to explore data involving this music. Curious people who love music would be inclined to use this.

# Setting up API

-   Go to <https://developer.spotify.com/documentation/web-api>

-   Log in to your Spotify account

-   Log into the dashboard (select "dashboard" )

-   Select the "Create an App" button

-   Enter the information necessary

-   Create the App

-   Go to Settings, view "Client Secret"

-   This will hold the Client ID and Secret needed to use Spotify API

# Packages

```{r}
#| include: false
library(tidyverse)
library(spotifyr)
library(knitr)
```

These packages will assist us in our search for the information we want.

# Accessing API in R

```{r}
#| eval: false
Sys.setenv(SPOTIFY_CLIENT_ID = 'YOUR_CLIENT_ID')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'YOUR_CLIENT_SECRET')
access_token <- get_spotify_access_token()
```

These lines of code call our API, and allows us to work with the data that we want to.

# Using Spotify API

I want to see what the most popular songs are for one of my favorite bands, Nirvana. Fortunately, we are able to do this using the data.

```{r}
get_artist_top_tracks(
  '6olE6TJLqED3rqDCT0FyPh', #nirvana ID
  market = "US",
  authorization = get_spotify_access_token(),
  include_meta_info = FALSE
) %>% select(name, popularity)
```

As we can see, their songs Smells Like Teen Spirit, Come As You Are, and Heart-Shaped Box are among the most popular Nirvana records.

With an awesome tool like this, we can find lots of information about all things music!
